---
title: "Group 21 CA pollution burden"
author: "Katie Pietrucha and Danny Feinberg"
date: "11/17/2021"
output: html_document
---

## Setup

```{r setup, include = TRUE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(janitor)
library(here)
library(hexbin)
```

## Read in & clean the data

```{r}
ces <- read_csv(here("data", "ces3demographics.csv")) %>% 
  clean_names()
```

## Visualize relationship between percentage of white residents and CalEnviroScreen 3.0 score

```{r}
ggplot(data = ces, aes(x = white_percent, y = ces_3_0_score)) +
  geom_hex() +
  theme_minimal() +
  scale_fill_gradient(low = "thistle", high = "midnightblue") +
  labs(y = "CalEnviroScreen 3.0 score",
       x = "Percent of Community that Identifies as White",
       caption = "Figure 1: The relationship between a community's percentage of White residents and its CalEnviroScreen 3.0 Score. Darker 
color indicates a higher density of observations. Data: OEHHA 2015.",
       fill = "Number of observations") +
  theme(legend.position = "bottom", plot.caption.position = "plot",
        plot.caption = element_text(hjust = 0))
```

It appears that communities that have a high percentage of residents that identify as White tend to have a lower CalEnviroScreen 3.0 score while communities with fewer White-identifying residents tend to have higher scores. This can be observed by the darker color (higher density) of observations in the lower right area and and higher left area of the graph. Therefore, there seems to be a negative correlation between the percentage of a community that identifies as White and CES 3.0 score total. As the "Whiteness" of a community decreases, the CES 3.0 score increases, indicating a higher pollution burden on Non-White communities.

## Citation 

Office of Environmental Health Hazard Assessment. (2018). Cal EnviroScreen 3.0. California Environmental Protection Agency. Retrieved November 16, 2021 from https://oehha.ca.gov/calenviroscreen/report/calenviroscreen-30.

## END TASK 2
