---
title: "Group 21 CA pollution burden"
author: "Katie Pietrucha and Danny Feinberg"
date: "11/17/2021"
output: html_document
---

## Setup

```{r setup, include = TRUE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(janitor)
library(here)
library(hexbin)
```

## Read in & clean the data

```{r}
ces <- read_csv(here("data", "ces3demographics.csv")) %>% 
  clean_names()
```

## Visualize relationship between percentage of white residents and CalEnviroScreen 3.0 score

```{r}
ggplot(data = ces, aes(x = white_percent, y = ces_3_0_score)) +
  geom_hex() +
  theme_minimal() +
  scale_fill_gradient(low = "thistle", high = "midnightblue") +
  labs(y = "CalEnviroScreen 3.0 score",
       x = "Percent of Community that Identifies as White",
       caption = "Figure 1: The relationship between a community's percentage of White residents and its CalEnviroScreen 3.0 Score. Darker 
color indicates a higher density of observations. Data: OEHHA 2015.",
       fill = "Number of observations") +
  theme(legend.position = "bottom", plot.caption.position = "plot",
        plot.caption = element_text(hjust = 0))
```

It appears that communities that identify primarily as white tend to have a lower CalEnviroScreen 3.0 score, which indicates low cumulative pollution burdens and a community that is less vulnerable to pollutants. This can be observed by the dark color (higher density) of observations in the lower right area of the graph. Conversely, it appears that communities that primarily identify as non-white tend to have higher CES 3.0 scores. Therefore, there seems to be a negative correlation between the percent of community that identifies as white and CES 3.0 score. As the "whiteness" of a community decreases, the CES 3.0 score increases, showing a higher pollution burden.

## Citation 

Office of Environmental Health Hazard Assessment. (2018). Cal EnviroScreen 3.0. California Environmental Protection Agency. Retrieved November 16, 2021 from https://oehha.ca.gov/calenviroscreen/report/calenviroscreen-30.

## END TASK 2
